### 2014-10-05 doc/tacs2014/article.tex
Final revision based on reviewer feedback

% Document title
% ==============
% Draft for conference/workshop paper to be submitted to http://tappcs.blogspot.mx/
% 2--4 pages in Spanish or English.

\documentclass[10pt,conference,a4paper]{IEEEtran}
\usepackage[utf8x]{inputenc}
\usepackage[hidelinks,pdftex]{hyperref}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage[T1]{fontenc}
\bibliographystyle{IEEEtran}

% This enables the kanji in the table, but all typefaces turn uglier.
% Also must be generated with xelatex
%\usepackage{xeCJK}
%\setCJKmainfont{cyberbit.ttf}

\hypersetup{
	pdftitle    = {A Modular Handwritten Kanji Recognition Schema},
	pdfauthor   = {Lars Fredrik Karlstr\"om and Everardo Guti\'errez L\'opez},
	pdfsubject  = {Workshop paper on the development of a modular design schema for handwritten kanji recognition.},
	pdfkeywords = {
		{Intelligent character recognition}
		{Handwriting recognition}
		{Kanji recognition}
		{Artificial neural network}
	},
}

\title{A Modular Handwritten Kanji Recognition Schema}

\author{
	\IEEEauthorblockN{Lars Fredrik Karlstr\"om}
	\IEEEauthorblockA{Faculty of Science, Dept. of Computer Science\\ Universidad Aut\'onoma de Baja California\\ \href{mailto:fredrik.karlstrm@uabc.edu.mx}{\texttt{fredrik.karlstrm@uabc.edu.mx}}}
	\and
	\IEEEauthorblockN{Everardo Guti\'errez L\'opez}
	\IEEEauthorblockA{Faculty of Science, Dept. of Computer Science\\ Universidad Aut\'onoma de Baja California\\ \href{mailto:everardo.gutierrez@uabc.edu.mx}{\texttt{everardo.gutierrez@uabc.edu.mx}}}
}

\pgfplotsset{
	compat = newest,
	xlabel near ticks,
	ylabel near ticks
}

\begin{document}
	\maketitle

	\begin{abstract}
		Handwriting continues to play a significant role in daily life, warranting continued research in the field of intelligent character recognition.
		East asian scripts, such as Japanese, remain challenging to process efficiently with high accuracy due to their vast number of symbols.
		In this work, a modular design schema based on reviews of previous work and the characteristics of the Japanese kanji sinographs is presented, along with
		the description of an ongoing effort to implement and evaluate a handwriting recognition system that incorporates online feature extraction,
		category clustering, and artificial neural networks.
	\end{abstract}

	\medskip
	\begin{IEEEkeywords}
		Intelligent character recognition, handwriting recognition, kanji recognition, artificial neural network.
	\end{IEEEkeywords}

	\section{Introduction}
	\label{sec:introduction}

	Although a variety of efficient digital input methods have become all but second nature to recent generations, for the vast majority
	handwriting remains an integral activity in daily life.
	The importance to further improve upon the state-of-the-art in handwriting recognition is underlined not only by the continued prevalence of 
	handwriting, but also by the need to digitalize archived material, its significant role in the development of fine motor skills,
	and a variety of other areas of application, such as signature verification \cite{plamondon2000online}.

	As an area of research, handwriting recognition predates even the digital computer. In 1888, Elisha Gray was awarded a patent %\cite{gray1888telautograph}
	for the Telautograph, which is considered a precursor to the stylus and digitizer, and in
	a 1957 publication, T. L. Dimond summarized early character recognition efforts and introduced the Stylator, an early input recognition device \cite{dimond1957devices}.

	The decades of research since has introduced a vast number of recognition methods for writing in any and all forms, with noteworthy successful techniques
	currently in widespread use including support vector machines, hidden Markov models, and artificial neural networks \cite{fujisawa2008forty, tappert1990state}.

	This paper documents a work in progress on the topic of intelligent handwriting recognition of Japanese characters.
	The remainder of the article is structured as follows:
	%In section \ref{sec:problem_description} we present the Japanese writing system and outline the primary issues with handwriting recognition for east asian script;
	%in section \ref{sec:previous_work}, we discuss some notable advances aimed at solving various aspects of the aforementioned problems.

	The Japanese writing system and an outline of the primary issues with handwriting recognition for east asian script is presented in section \ref{sec:problem_description}, and
	in section \ref{sec:previous_work}, some notable advances aimed at solving various aspects of the aforementioned problems are discussed.

	An explanation of the proposed solution approach is given in section \ref{sec:proposed_solution}, and an experimental implementation is described in section \ref{sec:experiments}
	along with some initial observations. Finally, planned future work is presented in section \ref{sec:future_work}.
	%whereafter we conclude by summarizing the contents of this paper in section \ref{sec:summary}.

	\begin{figure}
		\centering
		\includegraphics[width=2.75in]{./fig/yama-ryuu-kane.eps}
		\caption{Sample characters. The kanji ``yama'' (mountain), ``ry\=u'' (dragon), and ``kane'' (bell), written with 3, 10, and 20 strokes, respectively.}
		\label{fig_kanji_sample}
	\end{figure}

	\section{Problem description}
	\label{sec:problem_description}

	Any form of digital handwriting recognition is a complex undertaking,
	where the performance of the system is impacted by a multitude of factors---spanning
	everything from the quality and quantity of training data, to resilience toward input
	translation, rotation, and noise. As the number of candidate categories increase
	the size of the search space, however, the task quickly grows monumental.

	The Japanese writing system consists of the two \emph{kana} syllabaries \emph{hiragana} and \emph{katakana},
	as well as the sinographs commonly known as \emph{kanji}---literally ``Han characters''---that stem from China.
	The Japanese Ministry of Education's \mbox{official} ``regular use'' (j\=oy\=o) kanji list defines a set of 2,136
	baseline characters taught in elementary and secondary education \cite{hadamitzky2012japanese}.
	This set is far from exhaustive, though: the Japanese Industrial Standard X 0208 encoding, for instance, contains 6,355 different kanji.

	Although the characters are vast in number and differ significantly from one another in terms of complexity (see figures \ref{fig_kanji_sample}
	and \ref{fig_kanji_stroke_hist}), the writing system itself is not arbitrary. Any kanji may be broken down into smaller units, known as graphemes, 
	and they are commonly classified by basic, frequently occurring components referred to as radicals.
	Furthermore, stroke order is dictated by a set of rules. In essence, kanji are written from top to bottom and left to right within imaginary squares \cite{foerster1994kanji}.

	These properties are often leveraged in kanji recognition systems, but not without penalty: relying on the correct stroke order during online
	recognition, for instance, severely impacts recognition rates when a user provides erroneous input---even though it may still appear to be \emph{visually} correct \cite{shin2002optimal}.

	While certain aspects such as noise reduction and normalization strategies remain similar across systems, the aforementioned characteristics result in
	recognition systems designed for western input not performing satisfactorily when applied to east asian script \cite{jager2003state, tappert1990state}%, liu2004online}
	---unless the solution relies on complex techniques such as  ``deep'' neural networks \cite{ciresan2012multi}.

	%Given that the category search space for Japanese characters is an order of magnitude larger and contains a vastly differing set of
	%characteristics than its western counterparts, the same designs that yield highly accurate results on 
	%western handwriting recognition tasks do not perform satisfactorily when applied to asian script \cite{jager2003state, tappert1990state, liu2004online}
	%-- unless the solution relies on complex and computationally expensive techniques, such as  ``deep'' neural networks. \cite{ciresan2012multi}

	\begin{figure}
		\centering
		\begin{tikzpicture}[font=\small]
			\begin{axis}[
				ycomb,
				xlabel={Number of Strokes},
				ylabel={Number of Characters},
			]
				%\addplot[fill=white] coordinates {
				\addplot [] coordinates {
					(1, 6)
					(2, 32)
					(3, 64)
					(4, 115)
					(5, 151)
					(6, 199)
					(7, 318)
					(8, 434)
					(9, 473)
					(10, 544)
					(11, 581)
					(12, 581)
					(13, 532)
					(14, 435)
					(15, 434)
					(16, 369)
					(17, 299)
					(18, 207)
					(19, 175)
					(20, 135)
					(21, 106)
					(22, 69)
					(23, 49)
					(24, 37)
					(25, 20)
					(26, 12)
					(27, 8)
					(28, 6)
					(29, 3)
					(30, 2)
				};
			\end{axis}
		\end{tikzpicture}
		\caption{Distribution of 6,400 kanji by stroke count. \emph{Source: KanjiVG} \cite{kanjivg}.}
		\label{fig_kanji_stroke_hist}
	\end{figure}

	\section{Previous work}
	\label{sec:previous_work}

	A vast body of work has been amassed on the topic of Japanese handwriting recognition.
	This section provides a summary of publications dealing with many principal issues of interest,
	whose findings motivate this proposal.
	%Here we summarize some notable publications that influenced our adopted approach.

	In \cite{joe1989large}, Mori and Joe present selective and integrative neural networks that 
	segment categories into small groups, thereby improving the recognition rate of each network.

	After structuring the search space by clustering, Yang et al. perform two-stepped offline recognition by first
	comparing the input character to the centroids of each cluster, and then distinguishing between the members of the cluster
	whose centroid most closely matched the input \cite{yang2003accelerating}.

	In \emph{Hybrid Pen-Input Character Recognition System Based on Integration of Online-Offline Recognition},
	the authors present a system that leverages the advantages provided by each type of recognition.
	In addition to a discussion of various integration techniques, two feature extraction methods are described in detail: 
	a gradient-based technique for offline input, and a recursive substroke extractor for online input \cite{tanaka1999hybrid}.

	% Got A-H classification from here
	In \cite{nakai2001substroke}, the authors adapt a continuous speech recognition algorithm utilizing
	hidden Markov models for the purpose of online kanji recognition. Captured strokes are segmented into
	substrokes and classified based on directionality (see figure \ref{fig_stroke_categories});
	the resulting vector elements are thereafter employed as dictionary search keys.

	\emph{Online Handwritten Kanji Recognition Based on Inter-Stroke Grammar} explains the hierarchical structure of kanji,
	and explores how to leverage this information using stochastic context-free grammars \cite{ota2007online}.

	Finally, \emph{Best Practices for Convolutional Neural Networks Applied to Visual Document Analysis} by Simard, et al. outlines the
	potential of convolutional neural networks and provides a variety of techniques and guidelines directly applicable to the development of a fine-grain classifier \cite{simard2003best}.

	\begin{figure}
		\centering
		\includegraphics[width=2.75in]{./fig/stroke-categories.eps}
		\caption{A stroke may be classified according to its directionality.
			Further classes may be used to differentiate between long and short strokes,
		or to accomodate for pen-up motions. Here, the strokes of the kanji ``ki'' (tree) are identified as A-G-F-H.
		The dotted lines illustrate the pen-up motions that implicitly connect each successive stroke. Taking these motions into account,
		the label for the same kanji is A-4-G-3-F-2-H.}
		\label{fig_stroke_categories}
	\end{figure}

	\section{Proposed solution}
	\label{sec:proposed_solution}

	Taking the issues presented in section \ref{sec:problem_description} and the many influential works reviewed in section \ref{sec:previous_work}
	into consideration, an attempt is made to define a general description of a flexible, modular 
	design schema that splits the identification task into coarse and fine phases---thereby reducing the complexity of each undertaking---and 
	additionally permits replacing the normalization, extraction, segmentation, and classification strategies.
	The design schema is applicable to systems performing online as well as offline recognition tasks.
	Figure \ref{fig_model_overview} provides a high-level overview of the concept.

	\begin{figure}[b]
		\centering
		\includegraphics[width=2.75in]{./fig/model-overview.eps}
		\caption{The model consists of five components: input normalization methods, a feature extractor,
			a clustering strategy, a router (coarse classifier), and a fine-grained classifier.}
		\label{fig_model_overview}
	\end{figure}

	\subsection{Input normalization}

	Improving the input data, especially by making it more uniform with respect to available training data, has the potential
	to greatly increase recognition rates. In the case of offline input, techniques include noise reduction,
	cropping, and scaling; for online input, strokes may be analyzed and altered on-the-fly to increase consistency.
	Other possibilities include blurring or otherwise distorting input data in order to generalize it \cite{razzak2009combining}.

	\subsection{Feature extraction}

	The feature extraction algoritm should take online, offline, or combined input, and rapidly generate
	an n-dimensional vector representation of the provided data. The output should be sufficient for the
	routing component to reliably distinguish within which candidate subset the fine classifier should search.

	Generally speaking, a higher feature vector dimensionality would allow for the construction of more finely-tuned clusters,
	but this might well increase the likelihood of misrouting.%, resulting in a failed classification.

	Examples of feature extractors include recursive substroke segmentation and image gradation, both described in \cite{tanaka1999hybrid}.

	\subsection{Category clustering}

	By structuring the search space into clusters of similar \mbox{categories} based on extracted features,
	a corresponding number of fine classifiers that need only concern themselves with distinguishing
	between the symbols within their respective cluster may be constructed.
	Partitions could either be mutually exclusive or contain some degree of overlap.

	K-means clustering is often employed for this type of segmentation task, as is the Linde-Buzo-Gray algorithm.% \cite{yang2003accelerating}.
	%It is worth mentioning that optimally allocating $n$ elements in $k$ partitions is an NP-hard problem, even in the case of $k = 2$. \cite{drineas2004clustering}

	%This translates into significant reductions in model complexity,
	%which in turn implies both a decrease in training time
	%and the possibility of higher accuracy rates. Furthermore, given that no category overlap exists
	%between the clusters, each fine classifier may be trained in parallel, and each model may be gradually
	%improved as new training data is made available.

	%Centroid-based clustering
	%"Seeks to minimize the average squared distance between points in the same cluster"
	%goal is to choose k centers so as to minimize $\varphi$, the total squared distance between each point and its closest center.

	\subsection{Routing}

	Acting upon the vector provided by the feature extractor, the router---or coarse classifier---should indicate
	within which partition the category corresponding to the input is most likely located. his component could
	be contrived to indicate not only one but multiple targets in order to dampen the effect of erroneous routing.

	A 2014 publication by Zhu and Nakagawa documents the use of genetic algorithms to programmatically generate robust coarse classifiers \cite{zhu2014robust}.

	\subsection{Fine-grained classification}

	The final component of the system is the fine classifier, which is tasked with matching the
	given input against the candidates within its partition. As mentioned in section \ref{sec:previous_work}, a great
	variety of techniques exist, including support vector machines, Bayesian networks, hidden Markov models, and artificial neural networks.
	This step is not necessarily limited to using only one method; provided a reliable means of integration is developed,
	multiple classifiers could work in unity to provide the best possible classification.
	%As with the coarse classifier, this component could leverage either of or both online and offline information.

	%\section{Experiments \& preliminary results}
	\section{Experimental design \& preliminary results}
	\label{sec:experiments}

	An online kanji recognition system based on the design schema outlined in the previous section is
	currently being implemented. This system will be evaluated on established testing benchmarks
	as well as in the form of a mobile application. The diagram presented in figure \ref{fig_experimental_implementation} 
	explains how the various components interact.

	\begin{figure}
		\centering
		\includegraphics[width=2.75in]{./fig/experimental-implementation.eps}
		\caption{Experimental implementation. The input is first analyzed by the coarse classifier, resulting in the G-H-A\ldots feature representation.
		Thereafter, the routing network selects which cluster to search in, and the image data is passed to the corresponding CANN for fine-grain classification.}
		\label{fig_experimental_implementation}
	\end{figure}

	\subsection{Input normalization}

	Initially, two alternatives are being explored: an approach in which no input normalization is be applied, and one 
	in which the effect of rectifying online strokes by altering or discarding points below a certain threshold is evaluated.

	\subsection{Feature extraction}

	With the intent of progressively improving the various facets of the system,
	the first feature extractor was made as straightforward as possible: each stroke is directly
	classified with a letter, as illustrated in figure \ref{fig_stroke_categories}, based solely
	on its start and endpoints. Table \ref{tbl_stroke_analysis} presents a statistical overview of the
	result of classifying the 6,400 elements in the KanjiVG dataset in this manner.
	Intuitively, excluding pen-up motions should result in a more error-resilient system, due to
	the \emph{correct order} of the strokes not being taken into account.

	\begin{table}[b]
		\renewcommand{\arraystretch}{1.3}
		\caption{Stroke type frequency distribution overview }
		\label{tbl_stroke_analysis}
		\centering
		\begin{tabular}{ c | c c c c c l }
			\hline
			  & \bfseries \% & \small $\sum$ & $\mu$ & $\sigma$ & \footnotesize Highest occurrence \\ 
			\hline
			\hline
			\bfseries A & $33.20\%$ & $26,160$ & $4.09$ & $2.33$ & $15$ & \\%(蠶) \\
			\bfseries B & $1.47\%$  & $1,159$  & $0.18$ & $0.40$ & $2$  & \\%(摑) \\
			\bfseries C & $0.00\%$  & $0$      & --     & --     & --   & \\%     \\
			\bfseries D & $0.00\%$  & $0$      & --     & --     & --   & \\%     \\
			\bfseries E & $0.31\%$  & $245$    & $0.04$ & $0.19$ & $2$  & \\%(霪) \\
			\bfseries F & $12.72\%$ & $10,018$ & $1.57$ & $1.36$ & $10$ & \\%(鏐) \\
			\bfseries G & $29.98\%$ & $23,616$ & $3.69$ & $1.96$ & $13$ & \\%(鱸) \\
			\bfseries H & $22.32\%$ & $17,587$ & $2.75$ & $1.61$ & $11$ & \\%(靆) \\
			\hline
			            & $100.00\%$   & $78,785$ &        &        &      & \\
			\hline
		\end{tabular}
	\end{table}

	\subsection{Category clustering}

	Equipped with their respective feature vectors, the 6,400 categories were partitioned into 50
	mutually exclusive clusters using the K-means++ algorithm \cite{arthur2007k}.
	Based on the information provided
	in table \ref{tbl_stroke_analysis}, only the five significant dimensions A, B, F, G, and H were utilized as sorting parameters.
	The number of clusters was determined empirically, using Mardia's rule of thumb \mbox{$k \approx \sqrt{n / 2}$} \cite{mardia2005multivariate}
	as an upper limit.

	\subsection{Routing}

	Initially, three simple routing strategies are being applied: a lookup table, a function that
	identifies the most similar centroid, and a feedforward neural network. One auxiliary goal
	at this stage---in particular with the ANN---is creating some degree of resilience to errors
	such as a missing or incorrectly traced strokes.
	A differing aspect at this stage compared to previous works is the vastly decreased number
	of input features being considered, e.g. 8 versus 256 in \cite{yang2003accelerating}.

	\subsection{Fine-grained classification}

	Current efforts are focused on designing and evaluating convolutional neural networks to perform final, visual comparisons.
	\mbox{Convolutional} neural networks are known to provide a measure of resilience against translation and rotation,
	thus lending themselves well to offline character recognition tasks.

	The comparatively small number of category candidates per cluster (in current experiments averaging 128 kanji)
	implies that each model will require a lower degree of complexity. Furthermore, as it is probable that each
	partition will contain many similar-looking categories, it follows that the networks may be more likely to correctly
	discern between more minute differences that might have been generalized away in larger-scale topologies.
	These properties may lead to desireable reductions in execution time and storage space requirements,
	without considerable negative impact on recognition rates.

	%As previously stated, the comparatively small set of category candidates (in current experiments averaging 128 kanji)
	%implies that each cluster model requires a lower degree of complexity. It may also be expected that---seeing
	%how each partition is likely to contain many similar-looking characters---the networks will be more likely
	%to correctly discern between more minute differences that might have been generalized away in a larger-scale network.

	Finally, by focusing solely on offline character recognition at this stage, a significantly larger body of training data---such
	as the ETL character database, which contains hundreds of thousands of handwritten kanji samples \cite{yamamoto1986recognition}---may be utilized.

	\section{Future work}
	\label{sec:future_work}

	The experiments documented in this paper are an initial and ongoing attempt at implementing a kanji handwriting recognition \mbox{system}
	based on the outlined design methodology; numerous venues still remain to be explored.

	With regard to online recognition, for instance, the feature extraction strategy could be enhanced to break strokes into
	substrokes, diffentiate between long and short strokes, and take pen-up movements into account.
	This alteration would allow for clustering on up to 25 dimensions, rather than the currently available 8.
	Alternatively, by substituting the current coarse classifier with an offline feature extraction strategy, the entire system could
	be applied to exclusively offline handwriting recognition tasks, such as OCR of scanned documents.

	Additional clustering strategies could be examined, potentially harnessing the hierarchical properties of the symbols.
	Another venue would be developing a cascading clustering system with multiple coarse lookups, aimed at further
	reducing the size of the fine-grained classifier's search space.

	At the fine-grained recognition level, there is ample opportunity both to iterate upon the current convolutional neural network
	approach---refining the network topology and applying a variety of best practices \cite{simard2003best} in order to enhance the 
	recognition results---and to pursue other recognition techniques, such as hidden Markov models and support vector machines.

	% References
	\bibliography{./references}

\end{document}